---
title: "hw3"
format: pdf
editor: visual
---

# Homework 3 yay

Git repository: <https://github.com/cooperevan/ENVS-193DS_homework-03.git>

```{r, read in data, message=FALSE}
library(tidyverse)
library(here)
library(gt)
library(readxl)
library(janitor)
library(ggridges)
library(PNWColors)
```

## Problem 1: personal data

#### a. data summarization

The data could be summarized by comparing my observations of the response variable, hours asleep, to the type of workout I did on that day such as legs, push, or pull, to better understand if a different type of workout changes my sleep schedule, and can be completed by creating a boxplot. I can also compare sleep to other variables such as calories eaten that day, workout duration, time of workout, and screen time by creating a scatterplot of the data to compare.

#### **b. Visualization**

Using the summary you described in part a, **create a visualization of your data comparing your summarized response variable between groups**. If you are calculating a mean or median, show the underlying data in addition to your summary.

```{r}
STATS_workout_data_Sheet1_2 <- read_csv("~/Downloads/STATS workout data - Sheet1-2.csv")

workout_data <- STATS_workout_data_Sheet1_2
screen_data <- workout_data |> 
  clean_names() |> 
  drop_na() |> 
  rename(day = day_of_the_week) |>
  select(day, asleep, screen_time)


workout_clean <- workout_data |> 
  clean_names() |>
  rename(day = day_of_the_week, work_time = duration_of_workout) |>  
  select(day, asleep,exercise_type)

workout_01 <- workout_data |> 
  clean_names() |>
  rename(day = day_of_the_week, work_time = duration_of_workout) |>  
  select(day, asleep,exercise_type, work_time)

workout_legs <- workout_clean |> 
  filter(exercise_type == "Legs")
```

```{r, ggplots}
#affective visualization
ggplot(workout_clean, 
       aes(x = asleep, y = exercise_type, fill = exercise_type)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")+
   scale_fill_manual(values = PNWColors::pnw_palette("Bay", 5))+
  labs(x = "Sleep (Hours)",
       y = "Exercise",
       title = "Workout Data",
       caption = "Evan's Gym Plot")

ggplot(workout_clean,
       aes(x = exercise_type,
           y = asleep,
           color = exercise_type)) +
  geom_boxplot()+
  labs(x = "Exercise",
       y = "Time Asleep",
       title = "Workout Data",
       caption = "Evan's Gym Plot")

ggplot(workout_clean,
       aes(x = day,
           y = asleep,
           color = exercise_type))+
  geom_point()

ggplot(screen_data, aes(x = screen_time, 
                        y = asleep)) +
  geom_point(color = "darkslategray",
             size = 3,
             alpha = 0.7) +
  geom_smooth(method = "lm",
              se = TRUE,
              color = "tomato") +
  labs(
    title = "Relationship Between Screen Time and Sleep Time",
    caption = "evans data",
    x = "Screen Time (hours/day)",
    y = "Time Asleep (hours/day)") +
  theme_classic()

#plot with x = workout duration, y = sleep, and shape = type of workout
ggplot(workout_01, 
       aes(x = asleep, y = work_time, color = exercise_type))+
  geom_point(size = 3,
             alpha = 0.7)+
  theme()+
   scale_color_manual(values = PNWColors::pnw_palette("Shuksan2", 4))
```

Use colors in your figure (that are not the default ggplot settings).

Make sure your axis labels are full, readable text (not just your column names).

#### **c. Caption (5 points)**

Write a *caption* for your figure.

#### **d. Table presentation (10 points)**

Using `gt` or `flextable`, create a table with the same data summary that you describe in part a and visualized in part b. For example, if you described and visualized means, make a table with means. If you need to, round any numbers to one decimal point.

```{r}
workout_data |> gt()
```

## **Problem 2. Affective visualization (24 points)**

In this problem, you will create an **affective visualization** using your personal data in preparation for workshops during weeks 9 and 10.

In lecture, we talked about the three vertices of data visualization: 1) exploratory, 2) affective, and 3) communicative. We’ve done a lot of exploratory and communicative visualization, but have yet to think about affective visualization.

#### **a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences). (2 points)**

An affective visualization is a visualization that has provokes emotion in the viewer using creativity and novelty. For my personal data on whether type of workout impacts sleep, a possibility can be to use the website From Data to Viz to find interesting ways to display the data, like using the ggridge package and creating a ridge plot for the data.

#### **b. Create a sketch (on paper) of your idea.**

#### **c. Make a draft of your visualization.**

```{r}
ggplot(workout_clean, 
       aes(x = asleep, y = exercise_type, fill = exercise_type)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")+
   scale_fill_manual(values = PNWColors::pnw_palette("Bay", 5))+
  labs(x = "Sleep (Hours)",
       y = "Exercise",
       title = "Workout Data",
       caption = "Evan's Gym Plot")
```

#### **d. Write an artist statement. (8 points)**

An artist statement gives the audience context to understand your work. For each of the following points, write 1-2 sentences to address:

-   the content of your piece (what are you showing?)

    I'm showing the relationship between my sleep and type of workout

-   the influences (what did techniques/artists/etc. did you find influential in creating your work?)

-   the form of your work (written code, watercolor, oil painting, etc.)

-   your process (how did you create your work?)

## **Problem 3. Statistical critique**

#### **a. Revisit and summarize (6 points)**

What are the statistical tests the authors are using to address their main research question? (Note: you have already written about this in homework 2! Find that text and provide it again here!)

Insert the figure or table you described in Homework 2 here.

#### **b. Visual clarity (10 points)**

In 1-3 sentences, answer the question that best fits your paper.

If you inserted a figure in Part a: How clearly did the authors *visually* represent their statistics in figures? For example, are the x- and y-axes in a logical position? Do they show summary statistics (means and SE, for example) and/or model predictions, and if so, do they show the underlying data?

If you inserted a table in Part b: How clearly does the table represent the data underlying tests?

If you have neither: Critique another figure or table in the text for visual clarity (and insert a screenshot of that figure/table in your submission).
